---
output:
  word_document: default
  html_document: default
---
# Stephanie Thompson

## Module 4

### Classification Trees

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
library(RColorBrewer)
```

```{r}
library(readr)
Blood <- read_csv("C:/Users/thompsonsn/Downloads/Blood.csv")
View(Blood)
```

```{r}
library(readr)
parole <- read_csv("C:/Users/thompsonsn/Downloads/parole.csv")
View(parole)
```

```{r}
parolee = parole
```


```{r}
parolee= parolee %>% mutate(male = as_factor(as.character(male))) %>%
  mutate(male= fct_recode(male,
                          "female" = "0",
                          "male" = "1"))
parolee= parolee %>% mutate(race = as_factor(as.character(race))) %>%
  mutate(race= fct_recode(race, "NotWhite" = "2",
                          "White" = "1"))
```


```{r}
parolee= parolee %>% mutate(state = as_factor(as.character(state))) %>%
  mutate(state= fct_recode(state,
                           "Other" = "1",
                           "Kentucky" = "2", "Louisiana" = "3",
                           "Virginia" = "4"))
parolee= parolee %>% mutate(multiple.offenses = as_factor(as.character(multiple.offenses))) %>%
  mutate(multiple.offenses= fct_recode(multiple.offenses, "NotMultiple" = "0",
                                       "Multiple Offenses" = "1"))
parolee= parolee %>% mutate(crime = as_factor(as.character(crime))) %>%
  mutate(crime= fct_recode(crime,
                           "Other" = "1",
                           "larceny" = "2", "drug-related" = "3",
                           "driving-related" = "4"))
parolee= parolee %>% mutate(violator = as_factor(as.character(violator))) %>%
  mutate(violator= fct_recode(violator, "NoViolation" = "0",
                              "Violation" = "1"))

```

```{r}
str(parolee)
```

### 1.

```{r}
set.seed(12345)
train.rows = createDataPartition(y = parolee$violator, p=.7, list= FALSE)
train= parolee[train.rows,]
test= parolee[-train.rows,]

```



### 2.

```{r}
tree1 = rpart(violator ~., train, method = "class")
fancyRpartPlot(tree1)
```

### 3. If we wanted to estimate the outcome of a 40 year- old parolee from Louisiana who served 5 years, I would start at the top of the tree. Is the parolee from KY or VA, no so go right. Second tier, is the he younger than 43, yes so go left. Third tier, he did serve more than 2.6 years so left for yes. The outcome is No, he will not violate his parole. There is a 71% chance this is an accurate assumption.


### 4.

```{r}
plotcp(tree1)
```

```{r}
printcp(tree1)
```
### The cp for this tree is .054545




### 5.

```{r}
tree2 = prune(tree1, cp=tree1$cptable[which.min(tree1$cptable[,"xerror"]), "CP"])
table(train$violator)
```

### No violations is the majority with 418, while the parolees with violations is only 55.



### 6. 

```{r}
predtree = predict(tree1, train, type = "class")
head(predtree)
```

```{r}
confusionMatrix(predtree,train$violator,positive = "Violation")
```

### The accuracy or this tree is .907, Sensitivity is .491, and the Specificity is .962


### 7. 
```{r}
testpred =predict(tree1, newdata = test, type = "class")
confusionMatrix(testpred, test$violator, positive = "Violation")
```

### The Accuracy for this tree is .8614, the Sensitivity is .17391, and the Specificity is .95.



### 8.
```{r}
str(Blood)
```

```{r}
Blood= Blood %>% 
  mutate(DonatedMarch = as_factor(as.character(DonatedMarch))) %>%
  mutate(DonatedMarch= fct_recode(DonatedMarch,"No" = "0","Yes" = "1"))
```


### 9.

```{r}
set.seed(1234)
train.rows = createDataPartition(y = Blood$DonatedMarch, p=.7, list= FALSE)
train2= Blood[train.rows,]
test2= Blood[-train.rows,]
```

```{r}
bloodtree = rpart(DonatedMarch ~., train2, method = "class")
fancyRpartPlot(bloodtree)
```

```{r}
printcp(bloodtree)
```

### The best cp value for the training blood tree is .016


### 10.

```{r}
bloodtree2 = prune(bloodtree, cp=bloodtree$cptable[which.min(bloodtree$cptable[,"xerror"]), "CP"])
table(train2$DonatedMarch)
```

```{r}
bloodtestpred=predict(bloodtree2, newdata = test2, type = "class")
confusionMatrix(bloodtestpred, test2$DonatedMarch, positive = "Yes")
```


### The predictions here are not as good as with the Parole tree. Here is says 125 yes and 399 No. The accuracy here is only .7545. 